Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-07 paper/paper.tex
#22 edit down Levine in intro

The state system of early modern Europe was fraught with conflict.

This paper studies a different interaction between conflict and growth. I study how conflict affected the institutional quality of states. In particular, I investigate whether wars and conflict served as a selection mechanism - whether the constant competition caused better states to grow and worse states to shrink and eventually disappear.

These ideas are not new -

Institutional factors that benefit a state's economy also benefit a state's ability to win conflicts with others.

>>>>

\cite{levine2013, levine2021} argue that social evolution is driven by conflict. Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English. Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia. In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony. In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p. 439). strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.

How they define state power (pp. 439-440): "the potential ability of a social organisation to resist and influence other societies. This depends on institutional characteristics of the society, including stability of the government and the ability to collect taxes and conscript soldiers, which depend in turn on the law-abidingness of citizens, the efficiency of the courts, and the overall economic strength of the society." They *assume* in their model (p.443) that state power is directly proportional to free resources (resources in excess of subsistence).

I do not study their explanation for the European equilibrium that we observe in history; rather, I take a step back by taking the state system and the process of conflict-driven evolution as given. What interests me are the consequences of this process for economic growth.
Is it, as \cite{schoenholzer2022} argue explicitly, beneficial creative destruction - does state power equal state quality? Or are powerful states detrimental after all - because they engage in more extraction? \cite{tilly1985}

Why we would expect the hypothesis to be true: state power, defined as the ability to win conflicts with rivals, consists of factors like fiscal capacity, efficient courts, law-abiding citizens (\citealp{levine2021}, pp.439-449) - institutional factors that are also widely held to be beneficial for growth (see the vast literature on institutions). Add some institutions lit here.

My central identifying assumption: conquering a city implies more state power than inheriting it.

\subsection*{Other literature}
